Mining Interaction Log Data in a Creative Arts Practice using Transition Matrices

### 2015-09-24 flux-entropy-paper/flux-entropy-paper.tex
so close!

Our archive of interaction data-logs has been accumulated to support a
long-term, practice-led, research program into simple iPad interfaces
for group music-making similar to those described in Martin et
al~\cite{Martin:2014cr}. Over a period of more than two years (April
2013--June 2015), we collected data from group music-making sessions
where the participants used our custom-built, touch-screen apps on
Apple iPad devices. Although several different apps have been included
in our corpus, they shared a common mode of free-form touch
interaction where tapping produced short sounds and swirling or
swiping produced long sounds. Apart from these similarities, the apps
featured a variety of sound palettes, different kinds of visual
feedback, and different kinds of networked interactions (between the
iPads in an ensemble and with a central server).

In all of our 95 interaction sessions, every touch interaction---every
touch down, dragging movement and release---was time-stamped and
logged to disk for later analysis. These sessions included {\em rehearsals} for upcoming concerts, without an
audience, as well as curated {\em concerts}, in front of a live audience, where 
performative goals were paramount. They also included a number of {\em recording sessions}
for the purposes of formally-structured HCI experiments. In the recording
sessions, the performers were also interviewed and surveyed-- something which would have 
been disruptive and inappropriate for the concerts. Survey and interview data were not part of the
corpus. 
Some of the sessions were
iPad-only and others used iPads together with other acoustic
percussion instruments. Although the majority of the sessions were
free-form improvisations, a number of rehearsals and performances
involved composed pieces where the musicians read and performed
percussion notation using the apps.

Some of these session contexts are are shown
Figure~\ref{fig:three-performance-contexts} and other descriptive
statistics about the corpus are shown in Table \ref{corpus-table}.
In our corpus, the median length of sessions is 7m26s and the median
number of participants per ensemble is four. Each session record is
labelled with metadata about its \emph{context} (concert,
rehearsal, recording), its \emph{type} (improvisation or composition),
and its \emph{instrumentation} (iPads-only, or iPads with acoustic
percussion). The distribution of recorded sessions by context and type
is shown in Figure \ref{fig:count-data}.

@@

\begin{table}
\centering
\begin{tabular}{l|llll}
\hline
            & Total & Min  & Median   & Max     \\ 
\hline
N           & 95    &      &          &         \\
%Record size & 306MB     &         &        &   \\
Length & 13H9M19S & 1M19S & 7M26S & 22M20S \\
Participants& 353   & 2    & 4        & 9        \\
Flux &   & 0.02239 & 0.1445 & 0.3246\\
Entropy &   & 0.7683 & 3.342 & 4.442\\          
\hline
\end{tabular}
\caption{
  Representative statistics from the corpus of musical touch-screen
  interaction sessions described in this note.  Each record in the corpus
  consists of low level touch-data (such as touches begin, touches move, touches end) recorded in CSV format. 
  Many of the participants were
  present in multiple sessions, but this data was not tracked.\label{corpus-table}}
\end{table}

\section{Macro Gestures, Transition Matrices and Matrix Measures}

@@

\section{Quantitative transition matrix statistics}
\label{sec:underst-impr-group}

>>>>

Our archive of interaction data-logs has been accumulated to support a
long-term, practice-led, research program into simple iPad interfaces
for group music-making. Over a period of more than two years (April
2013--June 2015), we collected data from group music-making sessions
where the participants used our custom-built, touch-screen apps on
Apple iPad devices. Although several different apps have been included
in our corpus, they shared a common mode of free-form touch
interaction where tapping produced short sounds and swirling or
swiping produced long sounds. Apart from these similarities, the apps
featured a variety of sound palettes, different kinds of visual
feedback, and different kinds of networked interactions (between the
iPads in an ensemble and with a central server).

In all of our 95 interaction sessions, every touch interaction---
touches begin, touches move, touches end ---was time-stamped and
logged to disk for later analysis. These sessions included {\em rehearsals} without an
audience, as well as {\em concerts}, in front of a live audience, where 
performative goals were paramount. They also included a number of {\em
  recording} sessions
for the purposes of formally-structured HCI experiments. In the recording
sessions, the performers were also interviewed and surveyed-- something which would have 
been disruptive and inappropriate for the concerts. Survey and interview data were not part of the
corpus. 
Some of the sessions were
iPad-only and others used iPads together with other acoustic
percussion instruments. Although the majority of the sessions were
free-form improvisations, a number of rehearsals and performances
involved composed pieces where the musicians read and performed
percussion notation using the apps.
Some of these session contexts are are shown
Figure~\ref{fig:three-performance-contexts} and other descriptive
statistics about the corpus are shown in Table \ref{corpus-table}.
In our corpus, the median length of sessions is 7m26s and the median
number of participants per ensemble is four. Each session record is
labelled with metadata about its \emph{context} (concert,
rehearsal, recording), its \emph{type} (improvisation or composition),
and its \emph{instrumentation} (iPads-only, or iPads with acoustic
percussion). The distribution of recorded sessions by context and type
is shown in Figure \ref{fig:count-data}.

@@

\begin{table}
\centering
\begin{tabular}{l|llll}
\hline
            & Total & Min  & Median   & Max     \\ 
\hline
N           & 95    &      &          &         \\
%Record size & 306MB     &         &        &   \\
Length & 13H9M19S & 1M19S & 7M26S & 22M20S \\
Participants& 353   & 2    & 4        & 9        \\
Flux &   & 0.02239 & 0.1445 & 0.3246\\
Entropy &   & 0.7683 & 3.342 & 4.442\\          
\hline
\end{tabular}
\caption{
  Representative statistics from the corpus of musical touch-screen
  interaction sessions described in this note.  Each record in the corpus
  consists of low level touch-data recorded in CSV format. 
  Many of the participants were
  present in multiple sessions, but this data was not tracked.\label{corpus-table}}
\end{table}

\section{Macro Gestures and Transition Matrices}

@@

\section{Quantitative Transition Matrix Measures}
\label{sec:underst-impr-group}
